---
title: "Case Study: NYC Airbnb Data"
author: "Jessica Lavery"
date: "10/8/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# changing the size of a figure
# out.width/out.height control the size of the figure inserted into your document
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)
library(kableExtra)
library(leaflet)

# to apply the same formats/themes to all plots
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```


```{r}
# get airbnb dataset from p8105.datasets package
data(nyc_airbnb)

# look at structure of airbnb data
str(nyc_airbnb)

nyc_airbnb %>%
  count(room_type)

nyc_airbnb %>%
  count(neighbourhood_group)
```

Potential questions:  
1. How are airbnb prices related to rent in the neighborhood?  
2. Which neighborhoods have the most and least expensive rentals?  
3. Do hosts with multiple sites have higher prices or ratings than hosts with one site?
4. Does price have any relation to ratings?
5. Does price drive availability? 
6. Does room type correspond to ratings? Does this differ by neighborhoods?

### How are airbnb prices related to rent in the neighborhood?

Manhattan is the most expsnive borough, the Bronx is the least expensive. Staten Island has the greatest variability in price. 
```{r}
nyc_airbnb %>% 
  group_by(neighbourhood_group) %>% 
  summarize(median_price = median(price),
            mean_price = mean(price),
            sd_price = sd(price)) %>% 
  ggplot(aes(x = neighbourhood_group, y = mean_price)) + 
  geom_point() + 
  geom_errorbar(aes(x = neighbourhood_group, ymin = mean_price - sd_price, ymax = mean_price + sd_price)) + 
  labs(
    x = "Borough",
    y = "Average price (+/- standard deviation)"
  )
```

### Does price have any relation to ratings?

Higher prices are observed among airbnbs with higher reviews. 

```{r}
ggplot(data = nyc_airbnb %>% drop_na(review_scores_location), aes(x = as.factor(review_scores_location), y = price)) + 
  geom_boxplot()
```

### Is room type associated with star ratings?

```{r}
nyc_airbnb2 <- nyc_airbnb %>% 
  mutate(borough = neighbourhood_group,
         stars = factor(floor(review_scores_location/2))) %>% 
  drop_na(stars)

ggplot(nyc_airbnb2, aes(x = stars, y = price)) + 
  geom_boxplot()
```

### Distribution of price by room type
```{r}
ggplot(data = nyc_airbnb, aes(x = room_type, y = price)) + 
  geom_boxplot()
```

### Days booked by borough
```{r}
nyc_airbnb3 <- nyc_airbnb %>% 
  mutate(pct_booked = 100*((365 - availability_365)/365))
```

### Rating by number of times hosted
```{r}
# rating by number of times hosted
nyc_airbnb %>% 
  group_by(host_id) %>% 
  mutate(times_hosted = n()) %>% 
  ggplot(aes(x = factor(times_hosted), y = review_scores_location, color = factor(times_hosted))) + 
  geom_boxplot() + 
  labs(
    x = "Number of times hosted", 
    y = "rating"
  ) + 
  coord_flip()
```

### spatial locations

```{r}
nyc_airbnb %>% 
  filter(neighbourhood_group == "Manhattan", 
         price < 1000) %>% 
  ggplot(aes(x = lat, y = long, color = price)) + 
  geom_point(alpha = 0.3) +
  facet_grid(~room_type)
```

## Leaflet
```{r}
# creat stars variable
nyc_airbnb <- nyc_airbnb %>% 
  mutate(stars = floor(review_scores_location/2),
  label = paste0("<b>Stars: ", stars, "</b><br> Price: ", price))

# create color palet based on reviews
pal <- colorNumeric(
  palette = "viridis",
  domain = nyc_airbnb$stars)

nyc_airbnb %>% 
  filter(neighbourhood_group == "Manhattan") %>% 
  na.omit(stars) %>% 
  sample_n(1000) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  #grays out background
  addCircleMarkers(~lat, ~long, 
                   popup = ~label,
                   radius = 1,
                   color = ~pal(stars))
  # addMarkers(~lat, ~long, popup = ~name) #default markers
```

